Joint Likelihood Deconvolution of Astronomical Images in the Presence of Poisson Noise

### 2023-12-18 src/tex/ms.tex
[showyourwork] overleaf sync

    One possible approach to overcome this challenge has been proposed by \cite{Esch2004} and \cite{Connors2011} as the \enquote{Low Counts Image Reconstruction and Analysis} (LIRA\footnote{\dvd{LIRA was referred to as EMC2 in \cite{Esch2004}.}}) method. Their key idea is to introduce a prior assumption that favors smoothness on multiple spatial scales of the image. Instead of a trying to find a single point estimate for the reconstructed image they also sampled from the posterior distribution, which allowed them to quantify the associated uncertainties. However, the prior distribution is based on the generic assumption of smoothness and does not encode any \enquote{real world} or physical information on the structures present in the image.

    An interesting modification of the standard RL method is proposed by \cite{Ingaramo2014}. They show that RL can be effective in creating a merged image from multiple observations. They also observe that RL would take the best visual properties out of each image and merged them to create a more accurate estimate of the `ground truth'.

@@

    To account for both systematic effects we extend the model in Equation~\ref{eq:model} for the predicted counts, by a background normalisation factor $\alpha_j$ and a linear shift in position $\phi_j(\mathbf{x}| \delta_{x,j}, \delta_{y,j})$, which is specific for each observation. This yields the following modified model:
    %
    \begin{equation}
        \label{eq:model-npred-calibration}
        \mathbf{\lambda}_j = \mathbf{E}_j \cdot  \left(\mathrm{PSF}_j \circledast \phi_j(\mathbf{x}| \delta_{x,j}, \delta_{y,j})\right) + \alpha_j \cdot \mathbf{B}_j
    \end{equation}
    %
    In practice one observation is used as a reference and its corresponding systematic shifts in position $\delta_{x,j}, \delta_{y,j}$ are frozen during optimization. An example of the effect of the positional shift and background normalisation can be seen in Fig.~\ref{fig:npred-systematics}.

@@

    \begin{figure*}
        \centering\includegraphics{figures/scenarios.pdf}
        \caption{Illustration of the expected counts for the different source scenarios we used to generate test datasets. The source patterns shown are for the {\sl (A)}  {\tt points}, {\sl (B) {\tt asterism}}, {\sl (C)} {\tt shield}, and {\sl (D)} {\tt spiral} cases (see Section~\ref{sec:testmodels}). For the instrument scenario \chandra we convolve the ground truth image with a narrow Gaussian shaped PSF of width $\sigma=2$~pix. For the scenario \xmm we used a Gaussian shaped PSF with a width of $\sigma=2$~pix \dvd{same as \chandra?} and an effective area that is five times larger than that of \chandra. All images are of size $128{\times}128$ pixels.
        }
        \label{fig:scenarios}
        \script{scenarios.py}
    \end{figure*}

@@

        \item[\bf M2.] \textbf{\jolideco (Uniform, n=10)}: we use \jolideco with a uniform prior \dvd{on $\mathbf{x}$}, \dvd{running the optimizer for} 10 iterations. We consider this as a baseline method reference similar to RL. While the optimization procedure is different, we expect similar behaviour for the local correlations of pixels. This ad hoc approach of early stopping is often used in practice to regularise the RL output.

        \item[\bf M3.] \textbf{\jolideco (Uniform, n=1000)}:  we use \jolideco with a uniform prior \dvd{on $\mathbf{x}$}, but running \dvd{the optimzier} for 1000 iterations. This is the same as above but without early stopping. This gives rise to the issue of RL decomposing extended emission into point sources.

@@

    \subsection{Results}
    \subsubsection{Comparing Methods}
    %
    \begin{figure*}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/comparison-scenarios.pdf}
            \caption{
                Comparison of the different deconvolution methods and prior assumptions for selected source scenarios, as described in Section~\ref{subsec:test-datasets}. The simulations \dvd{depicted} %of the data for this figure 
                used a fixed background level of $\lambda_{Bkg} = \qty[mode = math]{0.01}{cts/pix}$, a uniform exposure, and a Gaussian PSF of width $\sigma_{PSF} = \qty[mode = math]{2}{pix}$, corresponding to the \enquote{Chandra} instrument scenario. The leftmost column shows the simulated counts and the second column %from the left 
                shows the underlying ground truth.% to compare. 
                The remaining columns show the reconstructed flux for each method and prior assumptions for \jolideconospace. To enhance weak structures the images use an $\arcsinh$ stretching with a scale parameter of $a=0.02$. The stretching is the same for all images. The methods are described in detail in Section~\ref{subsec:methods}. A more detailed representation of the results is available at \url{https://jolideco.github.io/jolideco-comparison}.
            }
            \label{fig:comparison-scenarios}
            \script{comparison-scenarios.py}
        \end{centering}
    \end{figure*}
    %
    First we compare the performance of the five methods (and their differing prior assumptions) on the different source scenarios. Figure \ref{fig:comparison-scenarios} shows the results for the selected subset of scenarios, i.e., on A1, B3, C3 and D4.  \dvd{What do the numbers refer to?}

    On scenario A1 all methods provide acceptable to excellent results for the isolated point sources in the upper right of the image. This demonstrates all methods take into account the PSF and background models correctly. The early stopped uniform prior of M2 shows an extension for these cases, because the optimization process has not finished. M3, which optimizes for 1000 iterations, clearly improves the result and shows almost perfect point sources corresponding to the size of approximately one pixel. The \jolideco method with the Zoran-Weiss  patch prior (M4) results in less sharp images, because the prior has been learned on natural images of landscapes and every day objects and scenes, which \dvd{typically} do not include point sources. The \jolideco method with the prior learned from GLEAM data (M5), which also includes point sources, shows much improved results \dvd{relative to M4 but is} 
    %However 
    not as sharp as the results obtained with the uniform prior (M3). In the lower left of the image the point sources are fainter and closer together, challenging all methods.% in succeeding to separate them into separate sources. 
    The uniform prior optimized \enquote{to end} (M3) shows good results, separating the sources well. All other methods, however, cannot fully resolve these weak and closely situated point sources.

@@

    The weak extended source with four embedded brighter point sources of Scenario C3 is challenging for all the methods. While the methods again reconstruct the point sources well, the visual differences are large for the weak underlying extended emission. The uniform prior with 1000 iterations decomposes the extended source into a grid of almost equidistant point sources resembling only the outer boundary of the original source. Pylira shows a similar tendency but improves the reconstruction of the extended source by leaving more emission in between the bright emission spots. The best results are achieved by the patch priors, which reconstruct a plausible extended source with approximately uniform brightness. The jet like features %in the same scenario 
    are decomposed into a \enquote{thread} of point sources by the uniform prior and Pylira method. The patch prior achieves a much more plausible reconstruction with almost uniform brightness for the linear structures. We also note that the patch prior does not show a preferred direction or bias in reconstruction, that could be possibly encoded in the training data. \dvd{Do you want to say something about the point soures?}

@@

    \subsubsection{Comparing Signal-to-Noise Scenarios}
    %
    \begin{figure*}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/comparison-signal-noise.pdf}
            \caption{
                Comparison of \jolideco with the GLEAM v0.1 prior for different background levels and signal strength, as represented by scenario D1-D5. The images are grouped by background level with each pair of columns showing the counts data in the left column and the reconstruction in the right column. The rows show the scenarios D1-D5 as described in Section~\ref{subsec:test-datasets}. The background levels correspond to \qty[mode = text]{0.001}{cts / pixel}, \qty[mode = text]{0.01}{cts / pixel} and \qty[mode = text]{0.1}{cts / pixel}. The complete results for the simulations and additional information are available at \url{https://jolideco.github.io/jolideco-comparison}.
            }
            \label{fig:comparison-signal-noise}
            \script{comparison-signal-noise.py}
        \end{centering}
    \end{figure*}
    %
    As a second experiment we explore the dependency of the results on the background level and signal strength. Figure~\ref{fig:comparison-signal-noise} shows the results for the \jolideco method with the GLEAM prior in Scenario D1-D5. The background level varies between $\lambda_{Bkg}= 0.01, 0.1$ and \qty[mode = text]{1}{counts/pixel}. As expected, the reconstruction quality improves with higher signal strength and lower background level. Visually the method converges to the ground truth for very high signal-to-noise ratio. Indeed with very high signal strength there is no dependency of the result on the background level, particularly in %This especially valid for 
    Scenario~D5. On the other hand, the method struggles to reconstruct a useful signal in the case of very high background (scenarios D1-3).

@@

    Using only the data from the \enquote{XMM} scenario results in lower quality reconstructions, due to the lower angular resolution. However, we find that the spiral arm structure is reconstructed at all levels of signal strength, even when barely visible by eye. The ring structure is sharpened considerably in all cases as well. The near-by point sources on the left and right of the spiral arm structure cannot be separated into individual point sources, but still appear with improved
    signal-to-noise ratio. This is an inherent limitation of the method.  \dvd{Limitation of the method or of the data?} \todo{how far are those apart compared to the PSF? Hoes does this compare to the point scenario with XMM?}

    Combining the data results in the best reconstruction quality. The addition of the lower resolution \enquote{XMM} data does not negatively affect the quality of the overall reconstruction. When comparing the D5 scenario we see that the \enquote{Chandra} reconstruction shows a very good reconstruction of the spiral structure but still shows speckles in the surrounding ring. The latter suggests that the prior is not strong enough to maintain the pixel to pixel correlations in the case of high signal strength. The strength of the prior could be increased with the $\beta$ parameter, but for this study we did not tweak individual parameters, but instead used the default configuration. The specklely behaviour is not observed in the \enquote{XMM} case, where the reconstruction of the ring is smoother. In this case the correlation between pixels that are further apart is stronger because of the larger correlation PSF. However, the reconstruction of the spiral structure in the center shows visual artifacts, with the individual arms not completely separated. The joint reconstruction shows the positive qualities of both individual reconstructions. This confirms the results from \cite{Ingaramo2014}, where the authors found that combining multiple measurements for deconvolution with RL to be a \textquote{general tool for combining images with complementary strengths}.

    \section{Application Examples}
    %
    \begin{figure*}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/example-chandra.pdf}
            \caption{
               \jolideco result for \chandra observations of SNR \textit{1E~0102.2–7219}. The bottom left image shows the summed counts from 25 observations. The bottom right image  shows the the \jolideco deconvolution. The flux is given as a counts rate relative to obsID 1308. The panels \textit{Zoom A}, \textit{Zoom B} and \textit{Zoom C} show a zoomed-in version of three example regions in the SNR, marked with the white rectangles in the large images. The colormap and scale used in the zoom insets is the same as used for the larger images.
            }
            \label{fig:example-chandra}
            \script{example-chandra.py}
        \end{centering}
    \end{figure*}
    %
    \subsection{Deep \chandra Observation}
    \subsubsection{Image Example}
    First we apply \jolideco to a set of short-exposure \chandra observations of the SNR \textit{1E~0102.2–7219} and combine them into a single long-exposure image using the patch prior learned from JWST data. In total, we chose 25 observations\footnote{Same list of observation IDs as for \url{https://chandra.si.edu/photo/2009/e0102/}} taken with  the ACIS-S instrument and no grating.

@@

    We assume a spatially constant background of \qty[mode = text]{1e-5}{cts/pixel} but allow the normalization to vary later in the reconstruction process. We assume the exposure to be spatially constant and express its absolute value for each observation as the ratio of observation time with respect to a reference observation. As reference we chose the % 
    obsID 1308. For \jolideco we use the GMM prior learned on the JWST Cas A image. As there are many counts in the image we additionally use an oversampling factor of 2. We allow the background level and position calibration to vary among the observations. As a reference for the positional shift we also use \dvd{obsID} 1308. Figure~\ref{fig:example-chandra} shows the final result. \jolideco clearly both improves the angular resolution and reduces the noise of the counts image.

    \textbf{Zoom A} of \dvd{Figure~\ref{fig:example-chandra}} details an example for a filament structure. \jolideco maintains the extended structure along the filament, while sharpening in the perpendicular direction. Looking at the upper right corner one can also see that weak background emission is suppressed, also improving the contrast of the image.

    \textbf{Zoom B} shows the region around the central compact object of the SNR. While the emission is barely visible in the counts image it is clearly \dvd{evident}
    %sticks 
    out above the background in the \dvd{\jolideco reconstruction.} %image reconstructed by \jolideconospace. 
    The remaining visible extension in the point source is encoded in the JWST patch prior, which also shows a finite extension for point sources. We note that, as illustrated in Section~\ref{subsec:methods} for Scenario~A, the apparent size of the point source \dvd{may be} %is 
    smaller if the source is brighter. \dvd{What does it mean for a point source to be smaller?} We estimate the position of point source by finding the brightest pixel in the map, obtaining $\alpha_{J2000}=01^h04^m02.75^s$ and $\delta_{J2000}=-72^d02^m00.19^s$ which is consistent with the value reported by \cite{Long2020}.

    \textbf{Zoom C} shows a circular filament structure of the SNR. In the counts data the structure is barely visible, while the \jolideco reconstruction % by \jolideco 
    shows a circular structure with a bright emission spot.

    Checking the residual images per observation (not shown here), we notice that the assumption of time independence of the source and the instrument response is not completely fulfilled in this case. %We saw 
    The detection efficiency of \chandra changed 
    \dvd{and the sub-structure in the remnant expanded over the time of} %between
    the observations 
    %and the sub-structure in the remnant expands over time. 
    Both effects are previously known and have been reported for example in \cite{Xi2019}. Despite the clear improvement in spatial resolution the proper motion of structures in the image might limit the quality of the result. The full analysis example is available at \url{https://github.com/jolideco/jolideco-chandra-examples}.

    \begin{figure*}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/chandra-e0102-zoom-a.pdf}
            \caption{
                The left panel shows the deconvolved flux and counts profile of the region illustrated in the right panel. The associated $3\sigma$ error band is shown transparent in the background. Both profiles are normalized such that they integrate to unity and are aligned such that the position of their peak coincides.
                \vlk{I notice that you avoid the bright knot to SW.  Any particular reason why?  You could legitimately exclude it to avoid pileup effects, but depends on the actual observed count rates.}
            }
            \label{fig:chandra-e0102-zoom-a}
            \script{chandra-e0102-zoom-a.py}
        \end{centering}
    \end{figure*}

    \subsubsection{Flux Profile}
    To \dvd{quantify} %get a more quantitative measurement of 
    the improvement in angular resolution \dvd{relative to what?} we compute a flux profile on the \dvd{\jolideco reconstruction} %image 
    of E0102. %reconstructed by \jolideconospace. 
    We chose a region perpendicular to the filament shown in Zoom~A of Figure~\ref{fig:example-chandra}. \dvd{We compute errors via the bootstrap, with resampled data obtained by sampling the event lists with replacement for each individual observation.}
    %To estimate the errors we use the bootstrapping method. For each individual observation we sample from the event list with replacement. 
    For each \dvd{resampled data set}, %realization of the set of observation 
    we run \jolideco and compute a flux profile on the reconstructed image. In this way we obtain a distribution of profiles, which we can use to compute an error band. For comparison we also compute the profiles and associated errors directly on the stacked bootstrapped counts \dvd{(without processing or fitting)}. The result is shown in Figure~\ref{fig:chandra-e0102-zoom-a}.

    We compute the full-width at half-max (FWHM) of each of the profiles, by fitting a Gaussian shape with a linear gradient as a baseline 
    for both the original counts and the \jolideconospace-deconvolved intensities. We find that the deconvolved profile is significantly narrower. For the counts profile, $\omega_C=0.77 \pm 0.22$~arcsec, \dvd{I don't think you have defined $\omega_C$.} where the error bar is obtained by computing the standard deviations over the bootstrap sample; and the deconvolved intensity profile has $\omega_J=0.52 \pm 0.03$~arcsec. \dvd{You say how you compute the error bar. How do you compute the central value? From the full original data of by averaging the bootstrap values?}

    Assuming that the PSF and the shape of %peak of the 
    profile are approximately Gaussian we expect %the following relation between 
    the width of the counts profile, the PSF, and \dvd{the \jolideconospace-reconstructed profile to be related as} %the flux profile:
    %
    \begin{equation}
        \omega_{J} = \sqrt{\omega_C^2 - \omega_{\mathrm{PSF}}^2}
    \end{equation}
    %
    With $\omega_{\mathrm{PSF}}\approx0.6$~arcsec\footnote{See projected radial profiles of the on-axis PSF at \url{https://space.mit.edu/cxc/marx/tests/PSF.html}.  The 1$\sigma$ radius is $\approx$0.26~arcsec (corresponding to 39\% enclosed counts in 2-D), suggesting that FWHM$_{{\rm PSF}}\approx{2.355{\cdot}0.26=0.62}$~arcsec.}, we expect a FWHM after deconvolution of $\omega_J\approx0.48$~arcsec. This is roughly consistent with the values we measured for the profile.

@@

    We chose the GMM patch prior learned from \chandra SNRs described in the Section~\ref{sssec:chandra-snrs}. Because of the limited statistics we do not oversampling but keep the pixel size of the counts image. 
    Figure~\ref{fig:example-fermi-lat} shows the data,  the \jolideco results, and corresponding residuals. The \fermi data is sparse, with %very 
    limited %number of 
    signal counts. \jolideco clearly improves the noise level as well as angular resolution of the data. \dvd{I don't think we can improve the noise of the data. I guess you mean the \jolideco-reconstruction is cleaner than the data?} The calibration parameters of the predicted counts converged to different norms for the background model and positional shifts of order of $<1$ pixel for the different event classes. In this way \jolideco corrects for systematic errors in the reconstruction of the position. One can see that the reconstruction includes background emission around Vela Junior, indicating that the Galactic diffuse background model is either underestimated in brightness and/or associated with systematic errors. The background estimate could be improved by choosing a larger region of interest. However this would also require a larger effort of modeling nearby sources, which we avoided here.

    To further check the quality of the reconstructed flux we compute a map of approximate residual significance \dvd{I would call these standardized residuals} defined by:

    \begin{equation}
        \label{eq:approx-sigma}
        \sigma \approx \frac{\hat{N}_{\mathrm{Data}} - \hat{N}_{\mathrm{Pred}}} {\sqrt{\hat{N}_{\mathrm{Pred}}}}
    \end{equation}

>>>>

    One possible approach to overcome this challenge has been proposed by \cite{Esch2004} and \cite{Connors2011} with the EMC2 and LIRA methods. Their key idea is to introduce a prior assumption that favors smoothness on multiple spatial scales of the image. Instead of a trying to find a single point estimate for the reconstructed image they also sampled from the posterior distribution, which allowed them to quantify the associated uncertainties. However, the prior distribution is based on the generic assumption of smoothness and does not encode any \enquote{real world} or physical information on the structures present in the image.

    An interesting modification of the standard RL method is proposed by \cite{Ingaramo2014}. They show that RL can be effective in creating a merged image from multiple observations. They also observe that RL would take the best visual properties out of each image and merged them to result in a reconstruction closer to the \enquote{ground truth}.

@@

    To account for both systematic effects we extend the model in Equation~\ref{eq:model} for the predicted counts, by a background normalisation factor $\alpha_j$ and a linear shift in position $\phi_j(\mathbf{x}| \delta_{x,j}, \delta_{y,j})$, which is specific for each observation. This yields the following modified model:
    %
    \begin{equation}
        \label{eq:model-npred-calibration}
        \mathbf{\lambda}_j = \mathbf{E}_j \cdot  \left(\mathrm{PSF}_j \circledast \phi_j(\mathbf{x}| \delta_{x,j}, \delta_{y,j})\right) + \alpha_j \cdot \mathbf{B}_j
    \end{equation}
    %
    In practice one observation is used as a reference and its corresponding systematic shifts in position $\delta_{x,j}, \delta_{y,j}$ are frozen during optimization. An example of the effect of the positional shift and background normalisation can be seen in Figure~\ref{fig:npred-systematics}.

@@

    \begin{figure*}
        \centering\includegraphics{figures/scenarios.pdf}
        \caption{Illustration of the expected counts for the different source scenarios we used to generate test datasets. The source patterns shown are for the {\sl (A)}  {\tt points}, {\sl (B) {\tt asterism}}, {\sl (C)} {\tt shield}, and {\sl (D)} {\tt spiral} cases (see Section~\ref{sec:testmodels}). For the instrument scenario \chandra we convolve the ground truth image with a narrow Gaussian shaped PSF of width $\sigma=2$~pix. For the scenario \xmm we used a Gaussian shaped PSF with a width of $\sigma=6$~pix and an effective area that is five times larger than that of \chandra. All images are of size $128{\times}128$ pixels.
        }
        \label{fig:scenarios}
        \script{scenarios.py}
    \end{figure*}

@@

        \item[\bf M2.] \textbf{\jolideco (Uniform, n=10)}: we use \jolideco with a uniform prior on $\mathbf{x}$ and running the optimizer for 10 iterations. We consider this as a baseline method reference similar to RL. While the optimization procedure is different, we expect similar behaviour for the local correlations of pixels. This ad hoc approach of early stopping is often used in practice to regularise the RL output.

        \item[\bf M3.] \textbf{\jolideco (Uniform, n=1000)}:  we use \jolideco with a uniform prior on $\mathbf{x}$, but running the optimzier for 1000 iterations. This is the same as above but without early stopping. This gives rise to the issue of RL decomposing extended emission into point sources.

@@

    \subsection{Results}
    \subsubsection{Comparing Methods}
    %
    \begin{figure*}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/comparison-scenarios.pdf}
            \caption{
                Comparison of the different deconvolution methods and prior assumptions for selected source scenarios, as described in Section~\ref{subsec:test-datasets}. The simulations depicted used a fixed background level of $\lambda_{Bkg} = \qty[mode = math]{0.01}{cts/pix}$, a uniform exposure, and a Gaussian PSF of width $\sigma_{PSF} = \qty[mode = math]{2}{pix}$, corresponding to the \enquote{Chandra} instrument scenario. The leftmost column shows the simulated counts and the second column shows the underlying ground truth. The remaining columns show the reconstructed flux for each method and prior assumptions for \jolideconospace. To enhance weak structures the images use an $\arcsinh$ stretching with a scale parameter of $a=0.02$. The stretching is the same for all images. The methods are described in detail in Section~\ref{subsec:methods}. A more detailed representation of the results is available at \url{https://jolideco.github.io/jolideco-comparison}.
            }
            \label{fig:comparison-scenarios}
            \script{comparison-scenarios.py}
        \end{centering}
    \end{figure*}
    %
    First we compare the performance of the five methods (and their differing prior assumptions) on the different source scenarios. Figure \ref{fig:comparison-scenarios} shows the results for the selected subset of scenarios, i.e., on A1, B3, C3 and D4.  \dvd{What do the numbers refer to?}

    On scenario A1 all methods provide acceptable to excellent results for the isolated point sources in the upper right of the image. This demonstrates all methods take into account the PSF and background models correctly. The early stopped uniform prior of M2 shows an extension for these cases, because the optimization process has not finished. M3, which optimizes for 1000 iterations, clearly improves the result and shows almost perfect point sources corresponding to the size of approximately one pixel. The \jolideco method with the Zoran-Weiss  patch prior (M4) results in less sharp images, because the prior has been learned on natural images of landscapes and every day objects and scenes, which typically do not include point sources. The \jolideco method with the prior learned from GLEAM data (M5), which also includes point sources, shows much improved results relative to M4 but is not as sharp as the results obtained with the uniform prior (M3). In the lower left of the image the point sources are fainter and closer together, challenging all methods. The uniform prior optimized \enquote{to end} (M3) shows good results, separating the sources well. All other methods, however, cannot fully resolve these weak and closely situated point sources.

@@

    The weak extended source with four embedded brighter point sources of Scenario C3 is challenging for all the methods. While the methods again reconstruct the point sources well, the visual differences are large for the weak underlying extended emission. The uniform prior with 1000 iterations decomposes the extended source into a grid of almost equidistant point sources resembling only the outer boundary of the original source. Pylira shows a similar tendency but improves the reconstruction of the extended source by leaving more emission in between the bright emission spots. The best results are achieved by the patch priors, which reconstruct a plausible extended source with approximately uniform brightness. The jet like features are decomposed into a \enquote{thread} of point sources by the uniform prior and Pylira method. The patch prior achieves a much more plausible reconstruction with almost uniform brightness for the linear structures. We also note that the patch prior does not show a preferred direction or bias in reconstruction, that could be possibly encoded in the training data. \dvd{Do you want to say something about the point soures?}

@@

    \subsubsection{Comparing Signal-to-Noise Scenarios}
    %
    \begin{figure*}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/comparison-signal-noise.pdf}
            \caption{
                Comparison of \jolideco with the GLEAM v0.1 prior for different background levels and signal strength, as represented by scenario D1-D5. The images are grouped by background level with each pair of columns showing the counts data in the left column and the reconstruction in the right column. The rows show the scenarios D1-D5 as described in Section~\ref{subsec:test-datasets}. The background levels correspond to \qty[mode = text]{0.001}{cts / pixel}, \qty[mode = text]{0.01}{cts / pixel} and \qty[mode = text]{0.1}{cts / pixel}. The complete results for the simulations and additional information are available at \url{https://jolideco.github.io/jolideco-comparison}.
            }
            \label{fig:comparison-signal-noise}
            \script{comparison-signal-noise.py}
        \end{centering}
    \end{figure*}
    %
    As a second experiment we explore the dependency of the results on the background level and signal strength. Figure~\ref{fig:comparison-signal-noise} shows the results for the \jolideco method with the GLEAM prior in Scenario D1-D5. The background level varies between $\lambda_{Bkg}= 0.01, 0.1$ and \qty[mode = text]{1}{counts/pixel}. As expected, the reconstruction quality improves with higher signal strength and lower background level. Visually the method converges to the ground truth for very high signal-to-noise ratio. Indeed with very high signal strength there is no dependency of the result on the background level, particularly in Scenario~D5. On the other hand, the method struggles to reconstruct a useful signal in the case of very high background (scenarios D1-3).

@@

    Using only the data from the \enquote{XMM} scenario results in lower quality reconstructions, due to the lower angular resolution. However, we find that the spiral arm structure is reconstructed at all levels of signal strength, even when barely visible by eye. The ring structure is sharpened considerably in all cases as well. The near-by point sources on the left and right of the spiral arm structure cannot be separated into individual point sources, but still appear with improved signal-to-noise ratio. This is an inherent limitation of the method.  \dvd{Limitation of the method or of the data?}

    Combining the data results in the best reconstruction quality. The addition of the lower resolution \enquote{XMM} data does not negatively affect the quality of the overall reconstruction. When comparing the D5 scenario we see that the \enquote{Chandra} reconstruction shows a very good reconstruction of the spiral structure but still shows speckles in the surrounding ring. The latter suggests that the prior is not strong enough to maintain the pixel to pixel correlations in the case of high signal strength. The strength of the prior could be increased with the $\beta$ parameter, but for this study we did not tweak individual parameters, but instead used the default configuration. The speckly behaviour is not observed in the \enquote{XMM} case, where the reconstruction of the ring is smoother. In this case the correlation between pixels that are further apart is stronger because of the larger correlation PSF. However, the reconstruction of the spiral structure in the center shows visual artifacts, with the individual arms not completely separated. The joint reconstruction shows the positive qualities of both individual reconstructions. This confirms the results from \cite{Ingaramo2014}, where the authors found that combining multiple measurements for deconvolution with RL to be a \textquote{general tool for combining images with complementary strengths}.

    \section{Application Examples}
    %
    \begin{figure*}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/example-chandra.pdf}
            \caption{
               \jolideco result for \chandra observations of SNR \textit{1E~0102.2–7219}. The bottom left image shows the summed counts from 25 observations. The bottom right image  shows the the \jolideco deconvolution. The flux is given as a counts rate relative to obsID 1308. The panels \textit{Zoom A}, \textit{Zoom B} and \textit{Zoom C} show a zoomed-in version of three example regions in the SNR, marked with the white rectangles in the large images. The colormap and scale used in the zoom insets is the same as used for the larger images.
            }
            \label{fig:example-chandra}
            \script{example-chandra.py}
        \end{centering}
    \end{figure*}
    %
    \subsection{Deep \chandra Observation}
    \label{sec:chandra-example}
    \subsubsection{Image Example}
    First we apply \jolideco to a set of short-exposure \chandra observations of the SNR \textit{1E~0102.2–7219} and combine them into a single long-exposure image using the patch prior learned from JWST data. In total, we chose 25 observations\footnote{Same list of observation IDs as for \url{https://chandra.si.edu/photo/2009/e0102/}} taken with  the ACIS-S instrument and no grating.

@@

    We assume a spatially constant background of \qty[mode = text]{1e-5}{cts/pixel} but allow the normalization to vary later in the reconstruction process. We assume the exposure to be spatially constant and express its absolute value for each observation as the ratio of observation time with respect to a reference observation. As reference we chose the % 
    obsID 1308. For \jolideco we use the GMM prior learned on the JWST Cas A image. As there are many counts in the image we additionally use an oversampling factor of 2. We allow the background level and position calibration to vary among the observations. As a reference for the positional shift we also use obsID 1308. Figure~\ref{fig:example-chandra} shows the final result. \jolideco clearly both improves the angular resolution and reduces the noise of the counts image.

    \textbf{Zoom A} of Figure~\ref{fig:example-chandra} details an example for a filament structure. \jolideco maintains the extended structure along the filament, while sharpening in the perpendicular direction. Looking at the upper right corner one can also see that weak background emission is suppressed, also improving the contrast of the image.

    \textbf{Zoom B} shows the region around the central compact object of the SNR. While the emission is barely visible in the counts image it is clearly evident
    out above the background in the \jolideco reconstruction. The remaining visible extension in the point source is encoded in the JWST patch prior, which also shows a finite extension for point sources. We note that, as illustrated in Section~\ref{subsec:methods} for Scenario~A, the remaining extension of the point source after deconvolution may be smaller if the source is brighter. We estimate the position of point source by finding the brightest pixel in the map, obtaining $\alpha_{J2000}=01^h04^m02.75^s$ and $\delta_{J2000}=-72^d02^m00.19^s$ which is consistent with the value reported by \cite{Long2020}.

    \textbf{Zoom C} shows a circular filament structure of the SNR. In the counts data the structure is barely visible, while the \jolideco reconstruction shows a distinct circular structure with a bright emission spot.

    Checking the residual images per observation (not shown here), we notice that the assumption of time independence of the source and the instrument response is not completely fulfilled in this case. The detection efficiency of \chandra changed and the sub-structure in the remnant expanded over the time of the observations. Both effects are previously known and have been reported for example in \cite{Xi2019}. Despite the clear improvement in spatial resolution the proper motion of structures in the image might limit the quality of the result. The full analysis example is available at \url{https://github.com/jolideco/jolideco-chandra-examples}.

    \begin{figure*}
        \begin{centering}
            \includegraphics[width=\linewidth]{figures/chandra-e0102-zoom-a.pdf}
            \caption{
                The left panel shows the deconvolved flux and counts profile of the region illustrated in the right panel. The associated $3\sigma$ error band is shown transparent in the background. Both profiles are normalized such that they integrate to unity and are aligned such that the position of their peak coincides.
                %\vlk{it is not being actively avoided -- the knot shows up after deconvolution.  But if referee asks, can say it also avoids pileup issues} \vlk{I notice that you avoid the bright knot to SW.  Any particular reason why?  You could legitimately exclude it to avoid pileup effects, but depends on the actual observed count rates.}
            }
            \label{fig:chandra-e0102-zoom-a}
            \script{chandra-e0102-zoom-a.py}
        \end{centering}
    \end{figure*}

    \subsubsection{Flux Profile}
    To quantify the improvement in angular resolution of the reconstructed image over the data we compute a flux profile on the \jolideco reconstruction of E0102. We chose a region perpendicular to the filament shown in Zoom~A of Figure~\ref{fig:example-chandra}. We compute errors via the bootstrap, with resampled data obtained by sampling the event lists with replacement for each individual observation. For each resampled data set we run \jolideco and compute a flux profile on the reconstructed image. In this way we obtain a distribution of profiles, which we can use to compute an error band. For comparison we also compute the profiles and associated errors directly on the stacked bootstrapped counts (without processing or fitting). The result is shown in Figure~\ref{fig:chandra-e0102-zoom-a}.

    We compute the full-width at half-max (FWHM) $\omega$ of each of the profiles, by fitting a Gaussian shape with a linear gradient as a baseline for both the original counts and the \jolideconospace-deconvolved intensities. In both cases we compute the mean of $\omega$. We find that the deconvolved profile is significantly narrower. For the counts profile, $\omega_C=0.77 \pm 0.22$~arcsec, where the error bar is obtained by computing the standard deviations over the bootstrap sample; and the deconvolved intensity profile has $\omega_J=0.52 \pm 0.03$~arcsec.

    Assuming that the PSF and the shape of profile are approximately Gaussian we expect
    the width of the counts profile, the PSF, and the \jolideconospace-reconstructed profile to be related as:
    %
    \begin{equation}
        \omega_{J} = \sqrt{\omega_C^2 - \omega_{\mathrm{PSF}}^2}
    \end{equation}
    %
    With $\omega_{\mathrm{PSF}}\approx0.6$~arcsec\footnote{See projected radial profiles of the on-axis PSF at \url{https://space.mit.edu/cxc/marx/tests/PSF.html}.  The 1$\sigma$ radius is $\approx$0.26~arcsec (corresponding to 39\% enclosed counts in 2-D), suggesting that FWHM$_{{\rm PSF}}\approx{2.355{\cdot}0.26=0.62}$~arcsec.}, we expect a FWHM after deconvolution of $\omega_J\approx0.48$~arcsec. This is roughly consistent with the values we measured for the profile.

@@

    We chose the GMM patch prior learned from \chandra SNRs described in the Section~\ref{sssec:chandra-snrs}. Because of the limited statistics we do not oversampling but keep the pixel size of the counts image. Figure~\ref{fig:example-fermi-lat} shows the data,  the \jolideco results, and corresponding residuals. The \fermi data is sparse, with limited signal counts. \jolideco clearly improves the noise level as well as angular resolution compared to the data. The calibration parameters of the predicted counts converged to different norms for the background model and positional shifts of order of $<1$ pixel for the different event classes. In this way \jolideco corrects for systematic errors in the reconstruction of the position. One can see that the reconstruction includes background emission around Vela Junior, indicating that the Galactic diffuse background model is either underestimated in brightness and/or associated with systematic errors. The background estimate could be improved by choosing a larger region of interest. However this would also require a larger effort of modeling nearby sources, which we avoided here.

    To further check the quality of the reconstructed flux we compute a map of standardized residuals defined by:

    \begin{equation}
        \label{eq:approx-sigma}
        \rho(N_{\mathrm{Data}}, N_{\mathrm{Pred}}) \approx \frac{\hat{N}_{\mathrm{Data}} - \hat{N}_{\mathrm{Pred}}} {\sqrt{\hat{N}_{\mathrm{Pred}}}}
    \end{equation}

    Where $\hat{N}$ is the weighted sum of counts from a Gaussian smoothing kernel with $\sigma_{\mathrm{smooth}}=5$ pixels. The result is shown in the right panel of Figure~\ref{fig:example-fermi-lat}. The image shows no systematic effects and a spatially homogeneous distribution of the residual noise. The resulting distribution of residual significances per pixel is centered at $\mu=-0.001$ with a width of $\sigma=0.991$, indicating a close agreement with a standard normal distribution, as expected for a good model of the emission. More details of the \fermi analysis are available at \url{https://github.com/jolideco/jolideco-fermi-examples}
